### 2015-02-27 icfp15-servo/servo.tex
Concurrency

\subsection{Libraries and abstractions}
Many high-level languages provide abstractions over I/O, threading, parallelism, and concurrency.
This area is one where we started out with much larger aspirations in Rust but ultimately had to pare back,
for reasons very similar to the reasons that we have avoided GC by default --- it is difficult to provide a
predictable, fast implementation that works across all platforms.
For example, thread-local storage (TLS) provides variables that are associated with a given thread.
In the presence of both lightweight (stackless) and full native threads across a variety of architectures, though,
it becomes difficult to understand what the cost of initialization and storage are for this feature.
Even lightweight threads in general, when interacting with a built-in scheduler and generic I/O system that
has to work across a variety of platforms (including Windows!), were too expensive to support in practice and
had to be removed.

Therefore, much like other modern browsers, Servo contains many of its own specialized implementations of
library functions that are tuned for the specific cases of web browsers.
For example, we have special ``small'' vectors that allow instantiation with a default inline size,
as there are use cases where we create many thousands of vectors, nearly none of which have more than 4~elements.
In that case, removing the extra pointer indirection --- particularly if the values are of less than pointer size ---
can be a significant space savings.
We also have our own work-stealing library that has been tuned to work on the flow tree during the process of layout,
as described in \secref{sec:browsers}.
It is our hope that this code might be useful to other projects as well, though it is fairly browser-specific today.

>>>>

\subsection{Libraries and abstractions}
Many high-level languages provide abstractions over I/O, threading, parallelism, and concurrency.
Rust provides functionality that addresses each of these concerns, but they are designed as thin
wrappers over the underlying services, in order to provide a predictable, fast implementation
that works across all platforms.
Therefore, much like other modern browsers, Servo contains many of its own specialized implementations of
library functions that are tuned for the specific cases of web browsers.
For example, we have special ``small'' vectors that allow instantiation with a default inline size,
as there are use cases where we create many thousands of vectors, nearly none of which have more than 4~elements.
In that case, removing the extra pointer indirection --- particularly if the values are of less than pointer size ---
can be a significant space savings.
We also have our own work-stealing library that has been tuned to work on the flow tree during the process of layout,
as described in \secref{sec:browsers}.
It is our hope that this code might be useful to other projects as well, though it is fairly browser-specific today.

Concurrency is available in Rust in the form of CML-style channels~\cite{reppy:cml}, but with a separation
between the reader and writer ends of the channel.
This separation allows Rust to enforce a multiple-writer, single-reader constraint, both simplifying and improving
the performance of the implementation over one that supports multiple readers.
We have structured the entire Servo browser engine as a series of threads that communicate over channels,
avoiding unsafe explicitly shared global memory for all but a single case (reading properties in the flow tree
from script, an operation whose performance is crucially tested in many browser benchmarks).
The major challenge we have encountered with this approach is the same one we have heard from other designers
of large concurrent systems --- reasoning about whether protocols make progress or threads eventually terminate
is manual and quite challenging, particularly in the presence of arbitrary thread failures.
